---
title: "AVENIR Mortality CONSORT Flow"
author: "Brittany Peterson"
date: "`r Sys.time()`"
output: 
  html_document:
    highlight: haddock
    theme: default
    code_folding: hide
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Overview

NOTE FOR PUBLIC REPLICATION
THIS SCRIPT WILL NOT RUN ON PUBLIC DATA

IT RELIES ON IDIOSYNCRATIC DATASETS TO IDENTIFY 91 GRAPPES THAT WERE RANDOMIZED BUT
NOT ACTUALLY MEASURED.  AFTER INTERNAL REVIEW, WE DECIDED TO
LEAVE AS-IS BECAUSE IT WOULD REQUIRE QUITE A BIT OF WORK TO WEAVE INTO THE PUBLIC
WORKFLOW AN APPROPRIATE DATASET TO IDENTIFY THESE GRAPPES
SEE THE NOTE IN FIGURE S1 CONSORT FLOW DIAGRAM TO EXPLAIN THIS
THIS IS WHY THERE ARE 2,909 GRAPPES WITH PERSON-TIME IN THE CHILD-PHASE LEVEL DATA

We have provided the original code and output for documentation

```{r setup, message = F, warning=F}
rm(list=ls())

library(here)

# source configuration file
source(here("R", "0-AVENIR-primary-Config.R"))

# load in PT by Grappe data
# rename public IDs to use code not run on public datasets
pt_by_grappe <- read_rds(here("data", "AVENIR_grappe_phase_public.rds")) %>%
  rename(siteid = siteid_public)

# load in PT by child for overall children
# rename public IDs to use code not run on public datasets
pt_by_child <- read_rds(here("data", "AVENIR_child_phase_public.rds")) %>%
  rename(siteid = siteid_public, hhid = hhid_public, childid = childid_public)

```     

```{r total randomized}

#-------------------------------------------
# load in randomization file
rando.groups <- read_csv(paste0(randomization_data, "/UNMASKED_AVENIR_grappe_enrollment_date_arm.csv"))

#-------------------------------------------
# load in unmasked treatment arms and merge to randomized files
tx_letter <- read_csv(paste0(randomization_data, "/avenir-treatment-assignments.csv")) %>%
    rename(tx = arm,
           treatment_letter = letter)

dosso_tx <- read_excel(paste0(randomization_data, "/AVENIR_Dosso_TxLetters_2023_03_08.xlsx"))
tahoua1_tx <- read_excel(paste0(randomization_data, "/AVENIR_Tahoua_TxLetters_0224_locked.xlsx"))
tahoua2_tx <- read_excel(paste0(randomization_data, "/AVENIR-Randomization-Tahoua-Part2-Mortality-2022-10-20_locked.xlsx"))

#-------------------------------------------
# clean randomization data
dosso_tx <- dosso_tx %>%
  rename(siteid = site_code) %>%
  filter(!is.na(treatment_letter_1_11)) %>%
  mutate(region = case_when(mortality_round4 == 1 ~ "Dosso Part2",
                            mortality == 1 ~ "Dosso Part1",
                            TRUE ~ "")) %>%
  select(region, siteid, treatment_letter_1_11)


tahoua1_tx <- tahoua1_tx %>%
  rename(siteid = site_code) %>%
  filter(mortality==1) %>%
  mutate(region = "Tahoua Part1") %>%
  select(region, siteid, treatment_letter_1_11)

tahoua2_tx <- tahoua2_tx %>%
  rename(siteid = site_code) %>%
  mutate(region = "Tahoua Part2") %>%
  select(region, siteid, treatment_letter_1_11)
tx <- bind_rows(dosso_tx, tahoua1_tx, tahoua2_tx)

#-------------------------------------------
#### get list of all grappes with random allocation and tx
tx_allocation_all <- tx %>%
    left_join(tx_letter, by = c("treatment_letter_1_11" = "treatment_letter")) %>%
    filter(region != "") %>%
    mutate(adaptive_rand = case_when(region == "Dosso Part1" ~ "Random Allocation 1",
                                     region == "Tahoua Part1" ~ "Random Allocation 2",
                                     region %in% c("Dosso Part2", "Tahoua Part2") ~ "Random Allocation 3")) %>%
    mutate(adaptive_rand = ifelse(siteid == "1-04-43-085", "Random Allocation 1", adaptive_rand))

total <- tx_allocation_all %>%
    group_by(adaptive_rand,tx) %>%
    summarise(rand = n())

#-------------------------------------------
# total siteids
total.site <- pt_by_child %>% distinct(siteid) %>% nrow()
total.child<- pt_by_child %>% distinct(childid) %>% nrow()
total.years <- sum(pt_by_grappe$pt_year_1to59)
total.deaths <- sum(pt_by_grappe$n_deaths_1to59)

#-------------------------------------------
# get # excluded
dosso_tx2 <- read_excel(paste0(randomization_data, "/AVENIR_Dosso_TxLetters_2023_03_08.xlsx")) %>%
    filter(mortality == 1 |mortality_round4 == 1)
tahouatx2 <- read_excel(paste0(randomization_data, "/AVENIR_Tahoua_TxLetters_02_01.xlsx")) %>%
    mutate(Menage = as.numeric(Menage),
           lat = as.numeric(lat),
           long = as.numeric(long),
           kmToTown = as.numeric(kmToTown)) %>%
    filter(mortality ==1 )
tx2 <- bind_rows(dosso_tx2, tahouatx2) %>%
  rename(siteid = site_code)

d.uncensused <- read_excel(paste0(final_data, "/final/AVENIR Refusals, Inexistent, and Low Coverage.xlsx"), sheet = 3) %>%
  rename(siteid = site_code)

not_in_pt <- tx2 %>%
    filter(!siteid %in% pt_by_child$siteid)

no.exist <- not_in_pt %>%
    filter(nonexistent == 1)

not_in_pt <- not_in_pt %>%
    left_join(d.uncensused, by = "siteid") %>%
    filter(is.na(nonexistent))

refus <- not_in_pt %>%
    filter(!siteid %in% no.exist$siteid) %>%
    filter(str_detect(Observation, "refus")) %>%
    distinct(siteid)

red_zone <- not_in_pt %>%
    filter(!siteid %in% no.exist$siteid) %>%
    filter(!siteid %in% refus$siteid) %>%
    filter(str_detect(Observation, "rouge")) %>%
    distinct(siteid)

other <- not_in_pt %>%
    filter(!str_detect(Observation, "rouge")) %>%
    filter(!str_detect(Observation, "refus"))

reasons <- tx2 %>%
    filter(!siteid %in% pt_by_child$siteid) %>%
    mutate(reason = case_when(siteid %in% no.exist$siteid ~ "Does not Exist",
                              siteid %in% refus$siteid ~ "Refus",
                              siteid %in% red_zone$siteid ~ "Red zone",
                              siteid %in% other$siteid ~ "other")) %>%
    group_by(reason) %>%
    summarise(n = n())
# BP note: 2-12-03-009 only had one household entered and did not contribute person time
# other 2 NA are protocol deviations

```    

# Last Census Date: July 31, 2023

# Grappe Randomization by Arm
```{r grappe rando by arm}

# table with 3 columns for arm, and one row for randomized, one row for analyzed
randomized <- tx_allocation_all %>%
    group_by(tx) %>%
    summarise(grappe = n()) %>%
    mutate(type = "Randomized") %>%
    spread(tx,grappe)

analyzed <- pt_by_grappe %>%
    filter(pt_year_1to59 > 0) %>%
    group_by(tx) %>%
    summarise(grappe = n_distinct(siteid)) %>%
    mutate(type = "Analyzed") %>%
    spread(tx,grappe)

grappe.totals <- bind_rows(randomized, analyzed)

kable(grappe.totals %>% rename(" " = type), "html",
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kable_styling()

```    

# Grappes Randomized and Analyzed
```{r grappe rando by arm 2}

# table with 3 columns for arm, and one row for randomized, one row for analyzed
randomized <- tx_allocation_all %>%
    group_by(adaptive_rand) %>%
    summarise(grappe = n()) %>%
    mutate(type = "Randomized") %>%
    spread(adaptive_rand,grappe)

analyzed <- pt_by_grappe %>%
    filter(pt_year_1to59 > 0) %>%
    group_by(adaptive_rand) %>%
    summarise(grappe = n_distinct(siteid)) %>%
    mutate(type = "Analyzed") %>%
    spread(adaptive_rand,grappe)

grappe.totals <- bind_rows(randomized, analyzed)

kable(grappe.totals %>% rename(" " = type), "html",
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kable_styling()

```    

# By Randomization Group

## Only children and grappes that contribute to person time for each individual phase are counted

```{r, message = F, warning=F}

# Random Allocation 1
total_child_a1 <- pt_by_child %>%
    filter(adaptive_rand == "Random Allocation 1") %>%
    distinct(childid) %>%
    nrow()

group1_flow <- pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 1") %>%
    group_by(phase) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names", values_to = "Values") %>%
    mutate(Values = round(Values,1))%>%
    bind_rows(., pt_by_grappe %>%
                  filter(adaptive_rand == "Random Allocation 1") %>%
                  summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = total_child_a1,
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
                  mutate(phase = "0-24") %>%
                  pivot_longer(1:5, names_to = "Names", values_to = "Values") %>%
                  mutate(Values = round(Values,1)))

# Random Allocation 2
total_child_a2 <- pt_by_child %>%
    filter(adaptive_rand == "Random Allocation 2") %>%
    distinct(childid) %>%
    nrow()

group2_flow <- pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 2") %>%
    group_by(phase) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names", values_to = "Values") %>%
    mutate(Values = round(Values,1)) %>%
    bind_rows(., pt_by_grappe %>%
                  filter(adaptive_rand == "Random Allocation 2") %>%
                  summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = total_child_a2,
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
                  mutate(phase = "0-24") %>%
                  pivot_longer(1:5, names_to = "Names", values_to = "Values") %>%
                  mutate(Values = round(Values,1))) %>%
    bind_rows(data.frame(phase = "6-12", `Number of Grappes` = 0, `Number of Children` = 0, `Person Years 1-59 months` = 0,
                         `Person Years 1-11 months` = 0, `Person Years 12-59 months` = 0, check.names = F) %>%
                  pivot_longer(2:6, names_to = "Names", values_to = "Values"), .) %>%
    bind_rows(data.frame(phase = "0-6", `Number of Grappes` = 0, `Number of Children` = 0, `Person Years 1-59 months` = 0,
                         `Person Years 1-11 months` = 0, `Person Years 12-59 months` = 0, check.names = F) %>%
                  pivot_longer(2:6, names_to = "Names", values_to = "Values"), .)

# Random Allocation 3

total_child_a3 <- pt_by_child %>%
    filter(adaptive_rand == "Random Allocation 3") %>%
    distinct(childid) %>%
    nrow()

group3_flow <- pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 3") %>%
    group_by(phase) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names", values_to = "Values") %>%
    mutate(Values = round(Values,1)) %>%
    bind_rows(., pt_by_grappe %>%
                  filter(adaptive_rand == "Random Allocation 3") %>%
                  summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = total_child_a3,
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
                  mutate(phase = "0-24") %>%
                  pivot_longer(1:5, names_to = "Names", values_to = "Values") %>%
                  mutate(Values = round(Values,1)))  %>%
    bind_rows(data.frame(phase = "12-18", `Number of Grappes` = 0, `Number of Children` = 0, `Person Years 1-59 months` = 0,
                         `Person Years 1-11 months` = 0, `Person Years 12-59 months` = 0, check.names = F) %>%
                  pivot_longer(2:6, names_to = "Names", values_to = "Values"), .) %>%
    bind_rows(data.frame(phase = "6-12", `Number of Grappes` = 0, `Number of Children` = 0, `Person Years 1-59 months` = 0,
                         `Person Years 1-11 months` = 0, `Person Years 12-59 months` = 0, check.names = F) %>%
                  pivot_longer(2:6, names_to = "Names", values_to = "Values"), .) %>%
    bind_rows(data.frame(phase = "0-6", `Number of Grappes` = 0, `Number of Children` = 0, `Person Years 1-59 months` = 0,
                         `Person Years 1-11 months` = 0, `Person Years 12-59 months` = 0, check.names = F) %>%
                  pivot_longer(2:6, names_to = "Names", values_to = "Values"), .)

# Overall

total_child_overall <- pt_by_child %>%
    distinct(childid) %>%
    nrow()

overall_flow <- pt_by_grappe %>%
    group_by(phase) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names", values_to = "Values") %>%
    mutate(Values = round(Values,1)) %>%
    bind_rows(., pt_by_grappe %>%
                  summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = total_child_overall,
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
                  mutate(phase = "0-24") %>%
                  pivot_longer(1:5, names_to = "Names", values_to = "Values") %>%
                  mutate(Values = round(Values,1)))


```     

## Random Allocation 1
```{r rando random allocation 1}

kable(group1_flow %>% select(-phase), "html", col.names = c("", ""),
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kableExtra::group_rows("Phase 0-6", 1, 5, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 6-12", 6, 10, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 12-18", 11, 15, label_row_css="l") %>%
  kableExtra::group_rows("Phase 18-24", 16, 20, label_row_css="l") %>%
  kableExtra::group_rows("Phase 0-24", 21, 25, label_row_css="l") %>%
  kable_styling()


```     

## Random Allocation 2
```{r rando allocation 2}

kable(group2_flow %>% select(-phase), "html", col.names = c("", ""),
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kableExtra::group_rows("Phase 0-6", 1, 5, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 6-12", 6, 10, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 12-18", 11, 15, label_row_css="l") %>%
  kableExtra::group_rows("Phase 18-24", 16, 20, label_row_css="l") %>%
  kableExtra::group_rows("Phase 0-24", 21, 25, label_row_css="l") %>%
  kable_styling()


```     

## Random Allocation 3
```{r rando Random Allocation 3}

kable(group3_flow %>% select(-phase), "html", col.names = c("", ""),
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kableExtra::group_rows("Phase 0-6", 1, 5, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 6-12", 6, 10, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 12-18", 11, 15, label_row_css="l") %>%
  kableExtra::group_rows("Phase 18-24", 16, 20, label_row_css="l") %>%
  kableExtra::group_rows("Phase 0-24", 21, 25, label_row_css="l") %>%
  kable_styling()


```     

## Overall Randomization Groups
```{r rando overall}

kable(overall_flow %>% select(-phase), "html", col.names = c("", ""),
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kableExtra::group_rows("Phase 0-6", 1, 5, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 6-12", 6, 10, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 12-18", 11, 15, label_row_css="l") %>%
  kableExtra::group_rows("Phase 18-24", 16, 20, label_row_css="l") %>%
  kableExtra::group_rows("Phase 0-24", 21, 25, label_row_css="l") %>%
  kable_styling()


```     

# By Randomization Group x Treatment Arm x Phase
```{r rand x treatment arm}
# 
# Random Allocation 1
total_child_a1 <- pt_by_child %>%
    filter(adaptive_rand == "Random Allocation 1") %>%
    distinct(childid) %>%
    nrow()

flow.arm.1 <- pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 1") %>%
    mutate(tx = case_when(
        tx == "3 - Placebo 1-59m" ~ "Placebo",
        tx == "2 - Azithro 1-11m" ~ "Azithro 1-11m",
        tx == "1 - Azithro 1-59m" ~ "Azithro 1-59m")) %>%
    group_by(phase,tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(3:7, names_to = "Names.1", values_to = "Values.1") %>%
        left_join(pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 2") %>%
    mutate(tx = case_when(
        tx == "3 - Placebo 1-59m" ~ "Placebo",
        tx == "2 - Azithro 1-11m" ~ "Azithro 1-11m",
        tx == "1 - Azithro 1-59m" ~ "Azithro 1-59m")) %>%
    group_by(phase,tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(3:7, names_to = "Names.2", values_to = "Values.2"),
    by = c("phase", "tx", "Names.1" = "Names.2")) %>%
        left_join(pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 3") %>%
    mutate(tx = case_when(
        tx == "3 - Placebo 1-59m" ~ "Placebo",
        tx == "2 - Azithro 1-11m" ~ "Azithro 1-11m",
        tx == "1 - Azithro 1-59m" ~ "Azithro 1-59m")) %>%
    group_by(phase,tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(3:7, names_to = "Names.3", values_to = "Values.3"),
    by = c("phase", "tx", "Names.1" = "Names.3")) %>%
    mutate(Values.1 = round(Values.1,1),
           Values.2 = round(Values.2,1),
           Values.3 = round(Values.3,1)) %>%
    replace_na(list(Values.2 = 0, Values.3 = 0))%>%
    ungroup()

overall.flow.arm <- pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 1") %>%
    mutate(tx = case_when(
        tx == "3 - Placebo 1-59m" ~ "Placebo",
        tx == "2 - Azithro 1-11m" ~ "Azithro 1-11m",
        tx == "1 - Azithro 1-59m" ~ "Azithro 1-59m")) %>%
    group_by(tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names.1", values_to = "Values.1") %>%
        left_join(pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 2") %>%
    mutate(tx = case_when(
        tx == "3 - Placebo 1-59m" ~ "Placebo",
        tx == "2 - Azithro 1-11m" ~ "Azithro 1-11m",
        tx == "1 - Azithro 1-59m" ~ "Azithro 1-59m")) %>%
    group_by(tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names.2", values_to = "Values.2"),
    by = c("tx", "Names.1" = "Names.2")) %>%
        left_join(pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 3") %>%
    mutate(tx = case_when(
        tx == "3 - Placebo 1-59m" ~ "Placebo",
        tx == "2 - Azithro 1-11m" ~ "Azithro 1-11m",
        tx == "1 - Azithro 1-59m" ~ "Azithro 1-59m")) %>%
    group_by(tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names.3", values_to = "Values.3"),
    by = c("tx", "Names.1" = "Names.3")) %>%
    mutate(Values.1 = round(Values.1,1),
           Values.2 = round(Values.2,1),
           Values.3 = round(Values.3,1)) %>%
    replace_na(list(Values.2 = 0, Values.3 = 0))%>%
    ungroup() %>%
    mutate(phase = "0-24")

flow.arm <- bind_rows(flow.arm.1, overall.flow.arm)

flow.arm$tx <- factor(flow.arm$tx, levels = c("Placebo","Azithro 1-11m","Azithro 1-59m"))
flow.arm$phase <- factor(flow.arm$phase, levels = c("0-6","6-12","12-18", "18-24", "0-24"))

flow.arm <- flow.arm %>%
    arrange(phase,tx)

                  
kable(flow.arm %>% select(-phase, -tx), "html", col.names = c("", "Random Allocation 1", "Random Allocation 2",
                                                         "Random Allocation 3"),
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kableExtra::group_rows("Phase 0-6", 1, 15, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 6-12", 16, 30, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 12-18", 31, 45, label_row_css="l") %>%
  kableExtra::group_rows("Phase 18-24", 46, 60, label_row_css="l") %>%
  kableExtra::group_rows("Phase 0-24", 61, 75, label_row_css="l") %>%
    pack_rows(index = c("Placebo" = 5, "Azithro 1-11 months" = 5, "Azithro 1-59 months" = 5,
                        "Placebo" = 5, "Azithro 1-11 months" = 5, "Azithro 1-59 months" = 5,
                        "Placebo" = 5, "Azithro 1-11 months" = 5, "Azithro 1-59 months" = 5,
                        "Placebo" = 5, "Azithro 1-11 months" = 5, "Azithro 1-59 months" = 5,
                        "Placebo" = 5, "Azithro 1-11 months" = 5, "Azithro 1-59 months" = 5),
              indent = T) %>%
  kable_styling()


```    

```{r overall phases and rando groups by arm}

overall.by.arm <- pt_by_grappe %>%
    group_by(tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names", values_to = "Values")


```    

  
# Session info
```{r session info}
sessionInfo()
```
