---
title: "AVENIR Mortality CONSORT Flow"
author: "Brittany Peterson"
date: "`r Sys.time()`"
output: 
  html_document:
    highlight: haddock
    theme: default
    code_folding: hide
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Overview

This file summarizes the number of clusters, children, and person time by treatment group. The numbers are reported in Figure S1 of the article. 

```{r setup, message = F, warning=F}
rm(list=ls())

library(here)

# source configuration file
source(here("R", "0-AVENIR-primary-Config.R"))

# load in PT by Grappe data
# rename public IDs to use code not run on public datasets
pt_by_grappe <- read_rds(here("data", "AVENIR_grappe_phase_public.rds")) %>%
  rename(siteid = siteid_public)

# load in PT by child for overall children
# rename public IDs to use code not run on public datasets
pt_by_child <- read_rds(here("data", "AVENIR_child_phase_public.rds")) %>%
  rename(siteid = siteid_public, hhid = hhid_public, childid = childid_public)

# this dataframe gives the reasons that some villages were not censused,
# number of villages not censused were calculated by comparing our data 
# with randomization documents and noting which villages do not appear in the data,
# reasons were given to us by the field team
# numbers are hard-coded as the document this comes from is not public
reason_not_censused <- structure(list(reason = c("Does not exist", "Other", "Red zone", 
"Refusal", "Protocol deviation"), n = c(39L, 2L, 40L, 7L, 3L)), row.names = c(NA, 
-5L), class = c("tbl_df", "tbl", "data.frame"))

```     

# Last Census Date: July 31, 2023

# Grappe Randomization by Arm
```{r grappe rando by arm}

# this dataframe gives the randomization numbers for each arm,
# number of villages in each arm were calculated via randomization documents
# numbers are hard-coded as the document this comes from is not public
randomized.tx <- structure(list(type = "Randomized", `1 - Azithro 1-59m` = 1273L, 
                                `2 - Azithro 1-11m` = 773L, `3 - Placebo 1-59m` = 954L), class = c("tbl_df", 
                                                                                                   "tbl", "data.frame"), row.names = c(NA, -1L))

analyzed <- pt_by_grappe %>%
    filter(pt_year_1to59 > 0) %>%
    group_by(tx) %>%
    summarise(grappe = n_distinct(siteid)) %>%
    mutate(type = "Analyzed") %>%
    spread(tx,grappe)

grappe.totals <- bind_rows(randomized.tx, analyzed)

kable(grappe.totals %>% rename(" " = type), "html",
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kable_styling()

```    

# Grappes Randomized and Analyzed
```{r grappe rando by arm 2}
# this dataframe gives the randomization numbers for each random allocation group,
# number of villages in each random allocation were calculated via randomization documents
# numbers are hard-coded as the document this comes from is not public
randomized.adaptive_rand <- structure(list(type = "Randomized", `Random Allocation 1` = 1407L, 
                                           `Random Allocation 2` = 1142L, `Random Allocation 3` = 451L), class = c("tbl_df", 
                                                                                                                   "tbl", "data.frame"), row.names = c(NA, -1L))

analyzed <- pt_by_grappe %>%
    filter(pt_year_1to59 > 0) %>%
    group_by(adaptive_rand) %>%
    summarise(grappe = n_distinct(siteid)) %>%
    mutate(type = "Analyzed") %>%
    spread(adaptive_rand,grappe)

grappe.totals <- bind_rows(randomized.adaptive_rand, analyzed)

kable(grappe.totals %>% rename(" " = type), "html",
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kable_styling()

```    

# By Randomization Group

## Only children and grappes that contribute to person time for each individual phase are counted

```{r, message = F, warning=F}

# Random Allocation 1
total_child_a1 <- pt_by_child %>%
    filter(adaptive_rand == "Random Allocation 1") %>%
    distinct(childid) %>%
    nrow()

group1_flow <- pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 1") %>%
    group_by(phase) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names", values_to = "Values") %>%
    mutate(Values = round(Values,1))%>%
    bind_rows(., pt_by_grappe %>%
                  filter(adaptive_rand == "Random Allocation 1") %>%
                  summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = total_child_a1,
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
                  mutate(phase = "0-24") %>%
                  pivot_longer(1:5, names_to = "Names", values_to = "Values") %>%
                  mutate(Values = round(Values,1)))

# Random Allocation 2
total_child_a2 <- pt_by_child %>%
    filter(adaptive_rand == "Random Allocation 2") %>%
    distinct(childid) %>%
    nrow()

group2_flow <- pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 2") %>%
    group_by(phase) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names", values_to = "Values") %>%
    mutate(Values = round(Values,1)) %>%
    bind_rows(., pt_by_grappe %>%
                  filter(adaptive_rand == "Random Allocation 2") %>%
                  summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = total_child_a2,
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
                  mutate(phase = "0-24") %>%
                  pivot_longer(1:5, names_to = "Names", values_to = "Values") %>%
                  mutate(Values = round(Values,1))) %>%
    bind_rows(data.frame(phase = "6-12", `Number of Grappes` = 0, `Number of Children` = 0, `Person Years 1-59 months` = 0,
                         `Person Years 1-11 months` = 0, `Person Years 12-59 months` = 0, check.names = F) %>%
                  pivot_longer(2:6, names_to = "Names", values_to = "Values"), .) %>%
    bind_rows(data.frame(phase = "0-6", `Number of Grappes` = 0, `Number of Children` = 0, `Person Years 1-59 months` = 0,
                         `Person Years 1-11 months` = 0, `Person Years 12-59 months` = 0, check.names = F) %>%
                  pivot_longer(2:6, names_to = "Names", values_to = "Values"), .)

# Random Allocation 3

total_child_a3 <- pt_by_child %>%
    filter(adaptive_rand == "Random Allocation 3") %>%
    distinct(childid) %>%
    nrow()

group3_flow <- pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 3") %>%
    group_by(phase) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names", values_to = "Values") %>%
    mutate(Values = round(Values,1)) %>%
    bind_rows(., pt_by_grappe %>%
                  filter(adaptive_rand == "Random Allocation 3") %>%
                  summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = total_child_a3,
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
                  mutate(phase = "0-24") %>%
                  pivot_longer(1:5, names_to = "Names", values_to = "Values") %>%
                  mutate(Values = round(Values,1)))  %>%
    bind_rows(data.frame(phase = "12-18", `Number of Grappes` = 0, `Number of Children` = 0, `Person Years 1-59 months` = 0,
                         `Person Years 1-11 months` = 0, `Person Years 12-59 months` = 0, check.names = F) %>%
                  pivot_longer(2:6, names_to = "Names", values_to = "Values"), .) %>%
    bind_rows(data.frame(phase = "6-12", `Number of Grappes` = 0, `Number of Children` = 0, `Person Years 1-59 months` = 0,
                         `Person Years 1-11 months` = 0, `Person Years 12-59 months` = 0, check.names = F) %>%
                  pivot_longer(2:6, names_to = "Names", values_to = "Values"), .) %>%
    bind_rows(data.frame(phase = "0-6", `Number of Grappes` = 0, `Number of Children` = 0, `Person Years 1-59 months` = 0,
                         `Person Years 1-11 months` = 0, `Person Years 12-59 months` = 0, check.names = F) %>%
                  pivot_longer(2:6, names_to = "Names", values_to = "Values"), .)

# Overall

total_child_overall <- pt_by_child %>%
    distinct(childid) %>%
    nrow()

overall_flow <- pt_by_grappe %>%
    group_by(phase) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names", values_to = "Values") %>%
    mutate(Values = round(Values,1)) %>%
    bind_rows(., pt_by_grappe %>%
                  summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = total_child_overall,
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
                  mutate(phase = "0-24") %>%
                  pivot_longer(1:5, names_to = "Names", values_to = "Values") %>%
                  mutate(Values = round(Values,1)))


```     

## Random Allocation 1
```{r rando random allocation 1}

kable(group1_flow %>% select(-phase), "html", col.names = c("", ""),
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kableExtra::group_rows("Phase 0-6", 1, 5, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 6-12", 6, 10, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 12-18", 11, 15, label_row_css="l") %>%
  kableExtra::group_rows("Phase 18-24", 16, 20, label_row_css="l") %>%
  kableExtra::group_rows("Phase 0-24", 21, 25, label_row_css="l") %>%
  kable_styling()


```     

## Random Allocation 2
```{r rando allocation 2}

kable(group2_flow %>% select(-phase), "html", col.names = c("", ""),
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kableExtra::group_rows("Phase 0-6", 1, 5, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 6-12", 6, 10, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 12-18", 11, 15, label_row_css="l") %>%
  kableExtra::group_rows("Phase 18-24", 16, 20, label_row_css="l") %>%
  kableExtra::group_rows("Phase 0-24", 21, 25, label_row_css="l") %>%
  kable_styling()


```     

## Random Allocation 3
```{r rando Random Allocation 3}

kable(group3_flow %>% select(-phase), "html", col.names = c("", ""),
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kableExtra::group_rows("Phase 0-6", 1, 5, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 6-12", 6, 10, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 12-18", 11, 15, label_row_css="l") %>%
  kableExtra::group_rows("Phase 18-24", 16, 20, label_row_css="l") %>%
  kableExtra::group_rows("Phase 0-24", 21, 25, label_row_css="l") %>%
  kable_styling()


```     

## Overall Randomization Groups
```{r rando overall}

kable(overall_flow %>% select(-phase), "html", col.names = c("", ""),
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kableExtra::group_rows("Phase 0-6", 1, 5, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 6-12", 6, 10, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 12-18", 11, 15, label_row_css="l") %>%
  kableExtra::group_rows("Phase 18-24", 16, 20, label_row_css="l") %>%
  kableExtra::group_rows("Phase 0-24", 21, 25, label_row_css="l") %>%
  kable_styling()


```     

# By Randomization Group x Treatment Arm x Phase
```{r rand x treatment arm}
# 
# Random Allocation 1
total_child_a1 <- pt_by_child %>%
    filter(adaptive_rand == "Random Allocation 1") %>%
    distinct(childid) %>%
    nrow()

flow.arm.1 <- pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 1") %>%
    mutate(tx = case_when(
        tx == "3 - Placebo 1-59m" ~ "Placebo",
        tx == "2 - Azithro 1-11m" ~ "Azithro 1-11m",
        tx == "1 - Azithro 1-59m" ~ "Azithro 1-59m")) %>%
    group_by(phase,tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(3:7, names_to = "Names.1", values_to = "Values.1") %>%
        left_join(pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 2") %>%
    mutate(tx = case_when(
        tx == "3 - Placebo 1-59m" ~ "Placebo",
        tx == "2 - Azithro 1-11m" ~ "Azithro 1-11m",
        tx == "1 - Azithro 1-59m" ~ "Azithro 1-59m")) %>%
    group_by(phase,tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(3:7, names_to = "Names.2", values_to = "Values.2"),
    by = c("phase", "tx", "Names.1" = "Names.2")) %>%
        left_join(pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 3") %>%
    mutate(tx = case_when(
        tx == "3 - Placebo 1-59m" ~ "Placebo",
        tx == "2 - Azithro 1-11m" ~ "Azithro 1-11m",
        tx == "1 - Azithro 1-59m" ~ "Azithro 1-59m")) %>%
    group_by(phase,tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(3:7, names_to = "Names.3", values_to = "Values.3"),
    by = c("phase", "tx", "Names.1" = "Names.3")) %>%
    mutate(Values.1 = round(Values.1,1),
           Values.2 = round(Values.2,1),
           Values.3 = round(Values.3,1)) %>%
    replace_na(list(Values.2 = 0, Values.3 = 0))%>%
    ungroup()

overall.flow.arm <- pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 1") %>%
    mutate(tx = case_when(
        tx == "3 - Placebo 1-59m" ~ "Placebo",
        tx == "2 - Azithro 1-11m" ~ "Azithro 1-11m",
        tx == "1 - Azithro 1-59m" ~ "Azithro 1-59m")) %>%
    group_by(tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names.1", values_to = "Values.1") %>%
        left_join(pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 2") %>%
    mutate(tx = case_when(
        tx == "3 - Placebo 1-59m" ~ "Placebo",
        tx == "2 - Azithro 1-11m" ~ "Azithro 1-11m",
        tx == "1 - Azithro 1-59m" ~ "Azithro 1-59m")) %>%
    group_by(tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names.2", values_to = "Values.2"),
    by = c("tx", "Names.1" = "Names.2")) %>%
        left_join(pt_by_grappe %>%
    filter(adaptive_rand == "Random Allocation 3") %>%
    mutate(tx = case_when(
        tx == "3 - Placebo 1-59m" ~ "Placebo",
        tx == "2 - Azithro 1-11m" ~ "Azithro 1-11m",
        tx == "1 - Azithro 1-59m" ~ "Azithro 1-59m")) %>%
    group_by(tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names.3", values_to = "Values.3"),
    by = c("tx", "Names.1" = "Names.3")) %>%
    mutate(Values.1 = round(Values.1,1),
           Values.2 = round(Values.2,1),
           Values.3 = round(Values.3,1)) %>%
    replace_na(list(Values.2 = 0, Values.3 = 0))%>%
    ungroup() %>%
    mutate(phase = "0-24")

flow.arm <- bind_rows(flow.arm.1, overall.flow.arm)

flow.arm$tx <- factor(flow.arm$tx, levels = c("Placebo","Azithro 1-11m","Azithro 1-59m"))
flow.arm$phase <- factor(flow.arm$phase, levels = c("0-6","6-12","12-18", "18-24", "0-24"))

flow.arm <- flow.arm %>%
    arrange(phase,tx)

                  
kable(flow.arm %>% select(-phase, -tx), "html", col.names = c("", "Random Allocation 1", "Random Allocation 2",
                                                         "Random Allocation 3"),
        booktabs=T, format.args = list(big.mark = ",")) %>%
  row_spec(0,bold=TRUE) %>%
  kableExtra::group_rows("Phase 0-6", 1, 15, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 6-12", 16, 30, label_row_css="l") %>% 
  kableExtra::group_rows("Phase 12-18", 31, 45, label_row_css="l") %>%
  kableExtra::group_rows("Phase 18-24", 46, 60, label_row_css="l") %>%
  kableExtra::group_rows("Phase 0-24", 61, 75, label_row_css="l") %>%
    pack_rows(index = c("Placebo" = 5, "Azithro 1-11 months" = 5, "Azithro 1-59 months" = 5,
                        "Placebo" = 5, "Azithro 1-11 months" = 5, "Azithro 1-59 months" = 5,
                        "Placebo" = 5, "Azithro 1-11 months" = 5, "Azithro 1-59 months" = 5,
                        "Placebo" = 5, "Azithro 1-11 months" = 5, "Azithro 1-59 months" = 5,
                        "Placebo" = 5, "Azithro 1-11 months" = 5, "Azithro 1-59 months" = 5),
              indent = T) %>%
  kable_styling()


```    

```{r overall phases and rando groups by arm}

overall.by.arm <- pt_by_grappe %>%
    group_by(tx) %>%
    summarise(`Number of Grappes` = n_distinct(siteid[pt_year_1to59>0]),
              `Number of Children` = sum(started_phase_1to59) - sum(lost_1to59),
              `Person Years 1-59 months` = sum(pt_year_1to59),
              `Person Years 1-11 months` = sum(pt_year_1to11),
              `Person Years 12-59 months` = sum(pt_year_12to59)) %>%
    pivot_longer(2:6, names_to = "Names", values_to = "Values")


```    

  
# Session info
```{r session info}
sessionInfo()
```
