---
title: "AVENIR Trial"
subtitle: "Cumulative enrollment and map"
author: "Ben Arnold F.I. Proctor Foundation (ben.arnold@ucsf.edu)"
date: "latest run: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: pygments
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---



# Summary

Map grappes and plot cumulative enrollment over time, colored by the three randomizations.

NOTE FOR PUBLIC REPLICATION: THIS SCRIPT REQUIRES GPS LOCATIONS OF STUDY COMMUNITIES, WHICH ARE CONSIDERED IDENTIFIABLE. THESE DATA ARE NOT AVAILABLE IN THE PUBLIC DATASETS

We have included this script and its output for documentation but it will only run on internal datasets.

# Load configuration file
```{r preamble, message=FALSE}
library(here)
here()
source(here("R","0-AVENIR-primary-Config.R"))
```


# Load data

Load the AVENIR trial analysis dataset and GPS coordinates

```{r load data}
#-------------------------------
# load the grappe enrollment dates
#-------------------------------
d_enrol <- read_csv(here(randomization_data,"UNMASKED_AVENIR_grappe_enrollment_date_arm.csv")) %>%
    rename(region2 = region)

#-------------------------------
# load the grappe gps coords
#-------------------------------
d_gps <- read_csv(here(final_data,"final/AVENIR_grappe_gps_2023-07-31.csv")) %>%
    rename(lat = median_latitude,
           lon = median_longitude)

#-------------------------------
# join gps to enrollment data
#-------------------------------
d1 <- d_gps %>%
    left_join(d_enrol, by=c("site_code"))

# confirm it was a clean 1:1 join
nrow(d1) == nrow(d_enrol)
nrow(d1) == nrow(d_gps)

# Recode tx labels to align with ms nomenclature
d2 <- d1 %>%
    mutate(
    tx = case_when(
        tx == "1 - Azithro 1-59m" ~ "Child Azithro",
        tx == "2 - Azithro 1-11m" ~ "Infant Azithro",
        tx == "3 - Placebo 1-59m" ~ "Placebo"
    ),
    tx = factor(tx, levels = c("Child Azithro","Infant Azithro","Placebo"))
           ) %>%
    # shorten adaptive randomization labels
    mutate(adaptive_rand = factor(adaptive_rand, 
                                  levels = c("Random Allocation 1", "Random Allocation 2", "Random Allocation 3"),
                                  labels = c("Randomization 1", "Randomization 2", "Randomization 3")))

```

# Cumulative enrollment by treatment and randomization phase

```{r tally cumulative enrollment by week, message = F, warning=F}

#-------------------------------
# sum the number of grappes
# enrolled in each week of the
# trial. Then get the cumulative
# sum
# do separately by randomization
# and tx
#-------------------------------
d_cumenrol <- d2 %>%
    # mutate(region2 = ifelse(site_code == "1-04-43-085", "Dosso Part1", region)) %>% # from Brittany, mis-categorized grappe
    mutate(start_date = as.Date(started_time, "%m/%d/%y"),
           min_date = min(start_date),
           enrol_week = floor(as.numeric(difftime(start_date,min_date, units = "weeks"))),
           enrol_week_ch = as.character(enrol_week)) %>%
    group_by(adaptive_rand,tx,enrol_week) %>%
    summarise(n = n()) %>%
    group_by(adaptive_rand,tx) %>%
    mutate(cumsum = cumsum(n)) 

#-------------------------------
# get the enrollment totals
# by randomization and tx
# match these to the max
# enrollment week in the data
# set n=0 (no new enrolled)
# for these observations
# as they are just placeholders
# for making the figure
#-------------------------------
d_enrol_totals <- d_cumenrol %>%
                  group_by(adaptive_rand,tx) %>%
                  summarise(cumsum = max(cumsum)) %>%
                  mutate(enrol_week = max(d_cumenrol$enrol_week), 
                         n=0)

# confirm cumsum totals match input data
sum(d_enrol_totals$cumsum) == nrow(d2)

#-------------------------------
# add the enrollment totals
# to a final observation in the
# data to get the cumulative
# totals to run through the
# end of the trial
#-------------------------------
d_cumenrol2 <- d_cumenrol %>%
    bind_rows(d_enrol_totals) %>%
    distinct() %>%
    arrange(adaptive_rand,tx,enrol_week) %>%
    # reverse the randomization factor for plotting
    mutate(adaptive_rand = factor(adaptive_rand, 
                                  levels = c("Randomization 3", "Randomization 2", "Randomization 1")))


#-------------------------------
# save the allocation times
# for each adaptation, to be
# used in the figure
#-------------------------------
adapt_week <- d_cumenrol2 %>% 
    group_by(adaptive_rand) %>% 
    summarise(adapt_time = min(enrol_week))

adapt_week

```

```{r plot cumulative grappe enrollment}
#-------------------------------
# plot cumulative enrollment
#-------------------------------
# pcols <- cbpal[c(8,7,6)]
pcols <- brewer_pal(palette = "YlOrRd")(9)[c(5,7,9)]

plot_cum_enrol <- ggplot(d_cumenrol2, aes(x = enrol_week, y = cumsum, fill = adaptive_rand)) +
    geom_area(alpha=0.9) +
    # geom_vline(xintercept = adapt_week$adapt_time[adapt_week$adaptive_rand=="Randomization 2"], 
    #            linetype = "solid", color = "grey20") +
    # geom_vline(xintercept = adapt_week$adapt_time[adapt_week$adaptive_rand=="Randomization 3"], 
    #            linetype = "solid", color = "grey20") +
    scale_x_continuous(breaks = seq(0,120,by=40)) +
    scale_y_continuous(breaks=seq(0,1200,by=200))+
    scale_fill_manual(values = pcols) +
    # scale_fill_brewer(type="seq") +
    xlab("Study week") +
    ylab("Cumulative communities enrolled") +
    theme_minimal() +
    facet_wrap(~tx) + 
    guides(fill=guide_legend(reverse = TRUE, title=""))  +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line()
    )

plot_cum_enrol
```  

# Randomization probabilities
```{r plot randomization probabilities}
#-------------------------------
# enter the adaptive randomization
# probabilities from Travis
#-------------------------------
rprobs <- data.frame(tx=rep(c("Child Azithro","Infant Azithro","Placebo"),3),
                     adaptive_rand = rep(c("Randomization 1","Randomization 2","Randomization 3"),c(3,3,3)),
                     rand = rep(c("1","2","3"),c(3,3,3)),
                     prob = c(0.333 ,0.333, 0.333, 
                              0.499, 0.204, 0.297,
                              0.554, 0.209, 0.237
                              )
                     ) %>%
    mutate(tx = factor(tx,levels=levels(d2$tx)),
           adaptive_rand = factor(adaptive_rand, levels=levels(d2$adaptive_rand)))

#-------------------------------
# make a bar plot of probabilities
#-------------------------------

plot_rprobs <- ggplot(data=rprobs, aes(x = rand, y = prob, fill = adaptive_rand)) + 
    facet_grid(~tx) +
    geom_bar(stat = "identity", color = NA, alpha = 1) +
    geom_text(aes(y = prob+0.04, label = prob), size=2) +
    scale_y_continuous(breaks=seq(0,0.6,by=0.2)) +
    scale_fill_manual(values = rev(pcols)) +
    labs(x = "Randomization", y = "Allocation prob.") +
    coord_cartesian(ylim=c(0,0.6)) +
    theme_minimal() +
    theme(
        legend.position = "none",
        panel.grid.minor.y = element_blank(),
        panel.grid.major = element_blank(), 
        axis.ticks.y = element_line()
        
    )

plot_rprobs
```

# Map grappes (communities)

```{r convert to spatial data}
#-------------------------------
# format grappe data as sf
#-------------------------------
d_sf <- st_as_sf(d2, coords=c("lon","lat"), 
                 crs=4326) # set coordinate ref system to WGS1984

```

```{r download admin shapes}

#-------------------------------
# load administrative boundaries
# downloaded from 
# humanitarian data exchange
# https://data.humdata.org/dataset/cod-ab-ner?
#-------------------------------

niger_adm0 <- st_read(dsn = here("data/ner_adm_shapefiles","ner_adm00_feb2018"))
niger_adm1 <- st_read(dsn = here("data/ner_adm_shapefiles","ner_adm01_feb2018"))
niger_adm2 <- st_read(dsn = here("data/ner_adm_shapefiles","ner_adm02_feb2018"))

# filter admin level 2 subdistricts to just Dosso and Tahoua
niger_adm2 <- niger_adm2 %>%
    filter(adm_01 %in% c("Dosso","Tahoua"))

#-------------------------------
# create a zoom window, and
# a cropped version of the admin
# polygons
#-------------------------------
sf_use_s2(FALSE)
niger_dists_cr <- st_crop(niger_adm1, xmin = 0, xmax = 7,
                                    ymin = 11, ymax = 19)

```

```{r map grappes}
# zoomed out overview of Niger
plot_niger_admin <- ggplot() +
    geom_sf(data=niger_adm1, color="black",lwd=0.5) +
    # geom_sf(data=niger_gadm0_sf, color="black",lwd=0.5) +
    coord_sf(crs = 4326) +
    geom_rect(aes(xmin = 0, xmax = 7, ymin = 11, ymax = 19), color = "black", fill = NA, lty = "dashed", inherit.aes = FALSE) +
    # add country labels
    annotate("text",x=4,y=22, label = "Niger", fontface="bold",size=8) +
    labs(x="",y="") +
    theme_minimal() +
    theme(
        panel.grid.major = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        plot.margin = margin(0,0,0,0, "pt")
        
          )
    
plot_niger_admin

# annotation line
dosso_lab_line <- data.frame(x1=4.45,x2=4.1,y1=12.5,y2=12.7)
tahoua_lab_line <-  data.frame(x1=5.9,x2=5.5,y1=17.9,y2=17.5)
    
    
# detailed zoom view, including grappes
plot_grappe_map <- ggplot() +
    # admin labels
    annotate("text",x=6,y=18.2, label = "Tahoua", fontface = "bold", color = "gray80", size=8) + # x=5.2,y=16.2,
    annotate("text",x=5.4,y=12.5, label = "Dosso", fontface = "bold", color = "gray80", size=8) +
    geom_segment(data=dosso_lab_line, aes(x=x1, xend=x2, y=y1, yend=y2), color = "gray60", lwd=0.8, arrow=arrow(angle=30,length=unit(2,"mm"),ends="last",type="open")) +
    geom_segment(data=tahoua_lab_line, aes(x=x1, xend=x2, y=y1, yend=y2), color = "gray60", lwd=0.8, arrow=arrow(angle=30,length=unit(2,"mm"),ends="last",type="open")) +
    # admin boundaries
    geom_sf(data=niger_adm2, fill=NA, color="black",lwd=0.3) +
    geom_sf(data=niger_adm1, fill=NA, color="black",lwd=1) + 
    # cluster points
    geom_sf(data=d_sf, aes(color=adaptive_rand),size=0.1, alpha=0.9, pch=19) +
    coord_sf(crs = 4326,
             xlim = c(0, 7), ylim = c(11, 19),
             expand = FALSE) +
    # scale_shape_manual(values=c(15,16,17)) +
    # scale_color_brewer(type="seq") +
    scale_color_manual(values=rev(pcols)) +
    annotation_scale(location = "br", width_hint = 0.3, pad_y=unit(0.1,"cm")) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.05, "in"), pad_y = unit(0.2, "in"),
        height=unit(0.75,"cm"), width=unit(0.75,"cm"),
        style = north_arrow_fancy_orienteering
        ) +
    labs(x="Longitude",y="Latitude") +
    guides(color=guide_legend(reverse = FALSE, title="", override.aes = list(size=4)))  +
    theme_minimal() +
    theme(
        legend.position = "bottom"
    )

plot_grappe_map

plot_map_with_inset <- plot_grappe_map + 
    inset_element(plot_niger_admin,
                  left = unit(0.12, "npc"), top = unit(1.35, "npc"),
                  bottom = unit(0.2, "npc"), right = unit(0.52, "npc"),
                  ignore_tag = TRUE
    )

# see patch for existing bug in patchwork for geom_sf
# https://stackoverflow.com/questions/18252827/increasing-area-around-plot-area-in-ggplot2
plot_map_with_inset$patches$layout$widths  <- 1
plot_map_with_inset$patches$layout$heights <- 1

plot_map_with_inset
```

# Composite figure

```{r make overall composite, fig.width=10, fig.height=10}

layout <- "
AB
AC
AC
"
patchwork_composite_fig1 <-  plot_map_with_inset  + plot_rprobs + plot_cum_enrol +
    plot_layout(design = layout, widths = c(0.7,0.3)) +
    plot_annotation(tag_levels = 'A') & 
    theme(plot.tag = element_text(size = 16, face = "bold"))

patchwork_composite_fig1

ggsave(filename = here("output","Figure-1-avenir-enrollment-map.pdf"), 
       patchwork_composite_fig1,
       device = "pdf", 
       width = 250,
       height = 160,
       units = "mm")
```

# Session Info
```{r session info}
sessionInfo()
```


