---
title: "AVENIR Trial"
subtitle: "Secondary analyses, all cause mortality"
author: "Ben Arnold F.I. Proctor Foundation (ben.arnold@ucsf.edu)"
date: "latest run: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: pygments
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---



# Summary

Select secondary analyses for AVENIR, all cause mortality

* Mortality rate among children ages 12-59 months, with rates compared between communities that receive azithromycin delivered only to children ages 1-11 months and communities that receive placebo

* Mortality rate among children ages 1-11 months, with rates compared between communities that receive azithromycin delivered only to children ages 1-11 months and communities that receive azithromycin delivered to children ages 1-59 months

# Load configuration file
```{r preamble, message=FALSE}
library(here)
here()
source(here("R","0-AVENIR-primary-Config.R"))
```


# Load data

Load the AVENIR trial analysis dataset

```{r load data}
#-------------------------------
# load the grappe-phase level 
# dataset
#-------------------------------
d1 <- read_rds(here("data","AVENIR_grappe_phase_public.rds")) %>%
  rename(siteid = siteid_public)

# d1$adaptive_rand[d1$siteid=="1-04-43-085"] <- "Random Allocation 1"
d2 <- d1 %>%
    mutate(
    tx = case_when(
        tx == "3 - Placebo 1-59m" ~ "Placebo",
        tx == "2 - Azithro 1-11m" ~ "Azithro 1-11m",
        tx == "1 - Azithro 1-59m" ~ "Azithro 1-59m"
    ),
    tx = factor(tx, levels = c("Placebo","Azithro 1-11m","Azithro 1-59m"))
           )

#-------------------------------
# pivot the data to be long
# by different age groups to
# make lots of the analyses
# and visualizations below easier
#-------------------------------
d2_long_ndeaths <- d2 %>%
    select(siteid,tx,adaptive_rand,district,phase, starts_with("n_deaths_")) %>%
    pivot_longer(cols = starts_with("n_deaths"), names_to = "age_group", names_prefix = "n_deaths_", values_to = "ndeaths")
d2_long_py<- d2 %>%
    select(siteid,tx,adaptive_rand,district,phase, starts_with("pt_year_")) %>%
    pivot_longer(cols = starts_with("pt_year_"), names_to = "age_group", names_prefix = "pt_year_", values_to = "py")
d2_long <- d2_long_ndeaths %>%
    left_join(d2_long_py, by = c("siteid","tx","adaptive_rand","district","phase","age_group")) %>%
    mutate(age_group = factor(age_group, levels=c("1to59","1to11","12to59")))

```

Confirm that number of deaths and person-time sum correctly for different age groups

```{r persontime death internal consistency check}
table(d2$n_deaths_12to59 + d2$n_deaths_1to11 == d2$n_deaths_1to59) # exactly equal because of integers
table(d2$pt_year_12to59 + d2$pt_year_1to11 - d2$pt_year_1to59 < 0.000001) # tol due to machine precision
```

## Collapse to grappe-level data

Sum over census intervals for grappe-level totals

These datasets are tailored for each between-arm comparison. 

In these secondary outcome analyses, there is a dataset for 12-59m mortality rates, including the Placebo and Azithro 1-11m groups, and a dataset with 1-11m mortality data, including the Azithro 1-11m and Azithro 1-59m groups. 

```{r collapse to grappe level datasets for analysis}
#-------------------------------
# summarize 12-59m deaths and pt 
# by grappe
# restrict to Placebo and 1-11m
#-------------------------------
# summarize to grappe-level data
# restrict to grappes with py>0 (drops 23 grappes)
table(d2$pt_year_12to59>0)
d12to59_cl <- d2 %>%
    filter(tx %in% c("Placebo","Azithro 1-11m") ) %>%
      group_by(siteid, adaptive_rand, tx) %>%
      summarize(ndeaths = sum(n_deaths_12to59),
                py = sum(pt_year_12to59), 
                .groups = "drop") %>%
    # limit to grappes with >0 person-years at risk
    filter(py>0)
table(d12to59_cl$tx)

#-------------------------------
# summarize 1-11m deaths and pt 
# by grappe
# restrict to 1-59m and 1-11m
#-------------------------------
# restrict to grappes with py>0 (drops 95 grappes)
table(d2$pt_year_1to11>0)
d1to11_cl <- d2 %>%
    filter(tx %in% c("Azithro 1-59m","Azithro 1-11m") ) %>%
      group_by(siteid, adaptive_rand, tx) %>%
      summarize(ndeaths = sum(n_deaths_1to11),
                py = sum(pt_year_1to11), 
                .groups = "drop") %>%
    # limit to grappes with >0 person-years at risk
    filter(py>0)
table(d1to11_cl$tx)

```

# Mortality rates

## Overall means

Summarize person-time, deaths, and mortality rates by grappe

```{r summarize person time deaths and mrate by grappe}
#-------------------------------
# summarize data at the grappe
# level, then pivot the data
# to long format to plot
#-------------------------------
dcl <- d2 %>%
    group_by(siteid, adaptive_rand, tx) %>%
    summarize(ndeaths1to59 = sum(n_deaths_1to59),
              pt1to59 = sum(pt_year_1to59), 
              ndeaths1to11 = sum(n_deaths_1to11),
              pt1to11 = sum(pt_year_1to11), 
              ndeaths12to59 = sum(n_deaths_12to59),
              pt12to59 = sum(pt_year_12to59), 
                .groups = "drop")
dcl_deaths <- dcl %>%
    select(siteid, adaptive_rand, tx, starts_with("ndeaths")) %>%
    pivot_longer(cols = starts_with("ndeaths"), names_to = "age_group", names_prefix = "ndeaths", values_to = "ndeaths")
dcl_pt <- dcl %>%
    select(siteid, adaptive_rand, tx, starts_with("pt")) %>%
    pivot_longer(cols = starts_with("pt"), names_to = "age_group", names_prefix = "pt", values_to = "py")
dcl2 <- dcl_deaths %>%
    left_join(dcl_pt, by = c("siteid","adaptive_rand","tx", "age_group")) %>%
    mutate(age_group = factor(age_group, levels=c("1to59","1to11","12to59")),
           mrate = ndeaths / py * 1000)

```

Summarize the mortality rates by age, treatment, and randomization group

Note: this is redundant with the estimates in the primary analysis, but repeated here as a quick internal consistency check to ensure the person-years and deaths match exactly.

```{r estimate mortality rates by tx group}
#-------------------------------
# summarize the mortality rates
# by treatment
#-------------------------------
d_mrates_tx <- d2_long %>%
    group_by(tx, age_group) %>%
    summarize(ndeaths = sum(ndeaths), py = sum(py), .groups = "drop") %>%
    mutate(mrate = 1000*ndeaths/py) %>%
    select(age_group, tx, ndeaths, py, mrate) %>%
    arrange(age_group, tx)

kbl(d_mrates_tx, digits = c(0,0,0,0,1), format.args = list(big.mark = ","),
    caption = "Mortality rates by treatment and age group") %>%
    kableExtra::kable_styling(bootstrap_options = "striped")

```

# Incidence Rate Ratios

```{r estimate IRRs}
#-------------------------------
# estimate the log IRR for 12-59m
# comparing arms 1-59m vs 1-11m
#
# estimate the IRR using poisson regression
# adjust for randomization group
#-------------------------------
fit12to59 <- glm(ndeaths ~ tx + adaptive_rand + offset(log(py)), family = "poisson", data=d12to59_cl )
summary(fit12to59)

#-------------------------------
# estimate the log IRR for 1-11m
# comparing arms 1-11m vs Placebo
#
# estimate the IRR using poisson regression
# adjust for randomization group
#-------------------------------
fit1to11 <- glm(ndeaths ~ tx + adaptive_rand + offset(log(py)), family = "poisson", data=d1to11_cl )
summary(fit1to11)


```

```{r bootstrap IRRs for CIs}

#-------------------------------
# bootstrap resample clusters
# to estimate 95% CIs for the IRR
# IRR for 12-59m
# comparing 1-11m vs Placebo
#------------------------------- 
breps <- 10000
irr12to59boot <- foreach(booti = 1:breps, .combine = rbind) %dopar% {
    #set seed within loop for perfect reproducibility using dopar
    set.seed(booti)
    di <- d12to59_cl[sample(1:nrow(d12to59_cl),replace=TRUE),]
    fiti <- glm(ndeaths ~ tx + adaptive_rand + offset(log(py)), family="poisson", data=di )
    return(fiti$coefficients[2])
}

( irr12to59ci <- quantile(irr12to59boot, probs=c(0.025, 0.975)) )


#-------------------------------
# bootstrap resample clusters
# to estimate 95% CIs for the IRR
# IRR for 1-11m
# comparing 1-59m vs 1-11m
#------------------------------- 
breps <- 10000
irr1to11boot <- foreach(booti = 1:breps, .combine = rbind) %dopar% {
    #set seed within loop for perfect reproducibility using dopar
    set.seed(booti)
    di <- d1to11_cl[sample(1:nrow(d1to11_cl),replace=TRUE),]
    fiti <- glm(ndeaths ~ tx + adaptive_rand + offset(log(py)), family="poisson", data=di )
    return(fiti$coefficients[2])
}

( irr1to11ci <- quantile(irr1to11boot, probs=c(0.025, 0.975)) )


```

```{r summarize IRRs}
#------------------------------- 
# combine IRRs into a data frame
# exponentiate logIRRs
#------------------------------- 
dIRRs <- data.frame(
    comparison = c("Azithro 1-11m vs. Placebo", "Azitrho 1-59m vs. Azithro 1-11m"),
    outcome = c("Mortality among 12-59m olds", "Mortality among 1-11m olds"),
    irr = exp(c(fit12to59$coefficients[2], fit1to11$coefficients[2])),
    irr_min95 = exp(c(irr12to59ci[1],irr1to11ci[1])),
    irr_max95 = exp(c(irr12to59ci[2],irr1to11ci[2]))
)

# print table
kbl(dIRRs, digits = 2, row.names = FALSE,
    caption = "Incidence Rate Ratio (IRR) estimates and 95% CIs, secondary between-group comparisons") %>%
    kableExtra::kable_styling(bootstrap_options = "striped")

```

```{r save IRRs for later use}
#-------------------------------
# save IRR estimates for later use
#-------------------------------
write_csv(dIRRs, file=here("output","avenir-mortality-secondary-IRRs.csv"))
write_rds(dIRRs, file=here("output","avenir-mortality-secondary-IRRs.rds"))
```

# Permutation tests

```{r conditional permutation function}
#-------------------------------
# conditional permutation function
# permute treatment within each 
# of 3 randomization strata
#
# this is very much tailored to our
# analysis dataset, but it is simple
#
# it stratifies the data.frame, 
# re-permutes treatment within strata,
# and then returns the data.frame with permuted treatment
#
# arguments:
# @x a grappe-level data frame with variables tx, adaptive_rand
#    where tx is the treatment label and adaptive_rand indicates randomization strata
#
# returns the same data frame, but with treatment shuffled
#-------------------------------
conditionally_permute <- function(x) {
    tx1 <- x[x$adaptive_rand == "Random Allocation 1",]
    tx2 <- x[x$adaptive_rand == "Random Allocation 2",]
    tx3 <- x[x$adaptive_rand == "Random Allocation 3",]
    tx1$tx <- sample(tx1$tx,size = nrow(tx1), replace=FALSE)
    tx2$tx <- sample(tx2$tx,size = nrow(tx2), replace=FALSE)
    tx3$tx <- sample(tx3$tx,size = nrow(tx3), replace=FALSE)
    return(bind_rows(tx1,tx2,tx3))
}

```


## Azithro 1-11m vs Placebo, Mortality 12-59m

```{r permutation test 12to59m}

# set a timer just to get a sense
start_time <- Sys.time()

npermute <- 10000
condp_null_12to59 <- foreach(permi = 1:npermute, .combine = rbind) %dopar% {
    
    # set seed within %dopar% for perfect replication
    set.seed(permi)

    #-------------------------------
    # permute the tx assignments
    # conditional on
    # randomization groups 1-3
    # sample w/o replacement within group
    #-------------------------------
    di <- conditionally_permute(d12to59_cl)
    
    #-------------------------------
    # fit a model
    # store the log IRR as the
    # test statistic
    #-------------------------------
    fiti <- glm(ndeaths ~ tx + adaptive_rand + offset(log(py)), family="poisson", data=di)
    logirr <- fiti$coefficients[2]
    return(data.frame(logirr=logirr))
    
}

# check time required
stop_time <- Sys.time()
difftime(stop_time,start_time)

#-------------------------------
# calculate the 2-sided 
# permutation test p-value
# the probability of observing
# a logIRR more extreme than
# that observed (absolute value 
# to make it 2-sided).
#-------------------------------
obs_irr_12to59 <- fit12to59$coefficients[2]
p_value_12to59 <- sum(abs(condp_null_12to59$logirr) >= abs(obs_irr_12to59)) / length(condp_null_12to59$logirr)
p_value_12to59_ci <- binom.test(sum(abs(condp_null_12to59$logirr) >= abs(obs_irr_12to59)), length(condp_null_12to59$logirr))$conf.int
cat("Azithro 1-11m v Placebo, Permutation P-value,",format(npermute,big.mark=","),"permutations:", round(p_value_12to59, 3),
    "\n                                                     Exact 95% CI:",round(p_value_12to59_ci[1],3),"to",round(p_value_12to59_ci[2],3), "\n")
    
```


```{r plot 12to59m permutation distribution}
#-------------------------------
# plot the null distribution 
# with the empirical value
# 1-11m vs Placebo
#-------------------------------
ggplot(condp_null_12to59, aes(x=logirr)) + 
  geom_histogram(alpha = 0.3, color = "gray20", bins = 50) +
  geom_vline(xintercept = obs_irr_12to59, color="black") +
  xlab("Null distribution of log(IRR) for 12-59 month mortality (Azithro 1-11m vs Placebo)") +
  theme_minimal()

```

## Azithro 1-59m vs Azithro 1-11m, Mortality 1-11m
```{r permutation test 1to11m}

# set a timer just to get a sense
start_time <- Sys.time()

npermute <- 10000
condp_null_1to11 <- foreach(permi = 1:npermute, .combine = rbind) %dopar% {
    
    # set seed within %dopar% for perfect replication
    set.seed(permi)

    #-------------------------------
    # permute the tx assignments
    # conditional on
    # randomization groups 1-3
    # sample w/o replacement within group
    #-------------------------------
    di <- conditionally_permute(d1to11_cl)
    
    #-------------------------------
    # fit a model
    # store the log IRR as the
    # test statistic
    #-------------------------------
    fiti <- glm(ndeaths ~ tx + adaptive_rand + offset(log(py)), family="poisson", data=di)
    logirr <- fiti$coefficients[2]
    return(data.frame(logirr=logirr))
    
}

# check time required
stop_time <- Sys.time()
difftime(stop_time,start_time)

#-------------------------------
# calculate the 2-sided 
# permutation test p-value
# the probability of observing
# a logIRR more extreme than
# that observed (absolute value 
# to make it 2-sided).
# get the exact binomial 95% CI
#-------------------------------
obs_irr_1to11 <- fit1to11$coefficients[2]
p_value_1to11 <- sum(abs(condp_null_1to11$logirr) >= abs(obs_irr_1to11)) / length(condp_null_1to11$logirr)
p_value_1to11_ci <- binom.test(sum(abs(condp_null_1to11$logirr) >= abs(obs_irr_1to11)), length(condp_null_1to11$logirr))$conf.int
cat("Azithro 1-59m vs Azithro 1-11m, Permutation P-value,",format(npermute,big.mark=","),"permutations:", round(p_value_1to11, 3),
    "\n                                                            Exact 95% CI:",round(p_value_1to11_ci[1],3),"to",round(p_value_1to11_ci[2],3), "\n")
    
```

```{r plot 1to11m permutation distribution}
#-------------------------------
# plot the null distribution 
# with the empirical value
# 1-11m vs Placebo
#-------------------------------
ggplot(condp_null_1to11, aes(x=logirr)) + 
  geom_histogram(alpha = 0.3, color = "gray20", bins = 50) +
  geom_vline(xintercept = obs_irr_1to11, color="black") +
  labs(x="Null distribution of log(IRR) for 1-11 month mortality (Azithro 1-59m vs Azithro 1-11m)",
       y="number of permutations") +
  theme_minimal()

```


# Session Info
```{r session info}
sessionInfo()
```


